---
title: "MacArthur Data"
author: "Kieran Healy"
date: "9/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rvest)
library(socviz)
library(ggrepel)
library(showtext)
showtext_auto()
library(myriad)
import_myriad_semi()

theme_set(theme_myriad_semi())
```

# Get a List of all MacArthur Fellows 

```{r}

urls <- paste0("https://www.macfound.org/fellows/", 1:1054, "/")



url <- "https://en.wikipedia.org/wiki/MacArthur_Fellows_Program"
page <- read_html(url) 
content <- html_node(page, "#mw-content-text")
years <- html_nodes(content, "h3")
winners <- html_nodes(page, ".column-width")
```

# Grab all their Fellow pages and put them in a list

Don't do this more than once! 

```{r, eval = FALSE}

## Grab the full Macfound page of every awardee. 
bio_pages <- urls %>% 
  map(~ {
    message(glue::glue("* parsing: {.x}"))
    Sys.sleep(5) # try to be polite
    safely(read_html)(.x)
  })

```


# Parse the pages


```{r}
get_fellow <- function(x) {
    ifelse(is.null(x), return(NA), FALSE)
  
    f_name <- rvest::html_node(x, ".has-top-margin") %>%
          rvest::html_text(trim = TRUE)
      
    f_year <- rvest::html_node(x, "h2") %>%
          rvest::html_text(trim = TRUE) %>%
          stringr::str_extract("Class of( January| February| March| April| May| June| July| August| September| October| November| December)?\\s+\\d{4}") %>%
          stringr::str_extract("\\d{4}")
  
    out <- rvest::html_node(x, ".photo-bio__content") %>%
          rvest::html_text(trim = TRUE) %>%
          stringr::str_remove_all("\n              ")
        
    f_title <- str_trim(str_match(out, "Title (.*?) Affiliation")[,2])
    f_affil <- str_trim(str_match(out, "Affiliation (.*?) Location")[,2])
    f_loc <- str_trim(str_match(out, "Location (.*?) Age")[,2])
    f_age <- str_trim(str_extract(out, "Age(.*?)\\d{2}")) %>% str_extract("\\d{2}")
    f_area <- str_trim(str_match(out, "Area of Focus (.*?)$")[,2])

    f_bio <- rvest::html_node(x, ":nth-child(3) .text-content--small") %>%
            rvest::html_text()     
    
    record <- tibble(
      name = f_name,
      year = f_year,
      title = f_title,
      affiliation = f_affil,
      location = f_loc,
      age = f_age,
      area = f_area,
      bio = f_bio)
    
    record
  }

```

```{r}

fellows <- bio_pages %>% 
  pluck("result") %>% # Get the webpages
  compact() %>% # Drop any empties
  map(get_fellow) %>% # Generate the individual records
  bind_rows() # and combine

fellows

```

# A bit more cleaning

```{r}

infer_gender <- function(bio){
  m_test <- str_detect(bio, "(\\b[Hh]e is)|(\\b[Hh]e was)|\\b[H]is\\b|\\b([Hh]e joined)|(\\b[Hh]e has)|\\b[Hh]e explore|\\b[Hh]e also|\\b[Hh]e produced|\\b[Hh]e founded|\\b[Hh]e worked|\\b[Hh]is work|\\b[Hh]is research")
  f_test <- str_detect(bio, "(\\b[Ss]he is)|(\\b[Ss]he was)|\\b[H]er\\b|\\b([Ss]he joined)|(\\b[Ss]he has)|\\b[Ss]he explore|\\b[Ss]he also|\\b[Ss]he produced|\\b[Ss]he founded|\\b[Ss]he worked|\\b[Hh]er work|\\b[Hh]er research")
  
  out <- case_when(
    m_test == TRUE & f_test == FALSE ~ "Male", 
    m_test == FALSE & f_test == TRUE ~ "Female",
    TRUE ~ NA_character_
    )
  out
}

fellows <- fellows %>%
  mutate(age = as.integer(age),
         year = int_to_year(as.integer(year)),
         sex = infer_gender(bio)) %>%
  select(name, year, age, sex, title, everything())

# tmp <- fellows %>% select(name, bio)
# tmp <- tmp %>%
#   mutate(sex = infer_gender(bio)) 
# 
# tmp %>% filter(is.na(sex)) %>% select(name)

## Eiko and Koma Otake
koma <- fellows %>% filter(name == "Eiko and Koma Otake")
koma$name <- "Koma Otake"

eiko <- fellows %>% filter(name == "Eiko and Koma Otake")
eiko$name <- "Eiko Otake"

fellows <- fellows %>% filter(name %nin% "Eiko and Koma Otake")
fellows <- rbind(fellows, eiko, koma)

## Manually code the remaining gender vals
fellows %>% filter(is.na(sex)) %>% select(name) %>% data.frame()

fellows <- fellows %>%
  mutate(sex = case_when(
  name == "Lin He" ~ as.character("Female"),
  name == "Aaron Shirley" ~ as.character("Male"),
  name == "Michael Lerner" ~ as.character("Male"),
  name == "Edward V. Roberts" ~ as.character("Male"),
  name == "Aaron Lansky" ~ as.character("Male"),
  name == "Patricia Locke" ~ as.character("Female"),
  name == "Robert H. McCabe" ~ as.character("Male"),
  name == "Jeraldyne Blunden" ~ as.character("Female"),
  name == "Martin Daniel Eakes" ~ as.character("Male"),
  name == "Louis Massiah" ~ as.character("Male"),
  name == "Eiko Otake" ~ as.character("Female"),
  name == "Koma Otake" ~ as.character("Male"),
  name == "Kun-Liang Guan" ~ as.character("Male"),
  name == "William W. McDonald" ~ as.character("Male"),
  name == "Juan Martin Maldacena" ~ as.character("Male"),
  name == "Ken Vandermark" ~ as.character("Male"),
  name == "Lucy Blake" ~ as.character("Female"),
  name == "Katherine Gottlieb" ~ as.character("Female"),
  name == "Nancy Siraisi" ~ as.character("Female"),
  name == "Whitfield Lovell" ~ as.character("Male"),
  name == "Junot Díaz" ~ as.character("Male"),
  name == "Jeffrey Brenner" ~ as.character("Male"),
  name == "Michelle Dorrance" ~ as.character("Female"),
  name == "Nicole Eisenman" ~ as.character("Female"),
  name == "LaToya Ruby Frazier" ~ as.character("Female"),
  name == "Taylor Mac" ~ as.character("Male"),
  #name == "Wu Tsang" ~ as.character(), ## one nonbinary case
  name == "Cameron Rowland" ~ as.character("Male"),
  TRUE ~ sex))

```

# Clean the Area field a little

```{r}

clean_area <- function(x) {
  stringr::str_remove(x, "(Website|Twitter).*$") %>%
  stringr::str_squish()
}

fellows <- fellows %>%
  mutate(area = clean_area(area))

```

# Clean Affiliation and Title

```{r}
fellows <- fellows %>%
  mutate(affiliation = replace_na(affiliation, "None Provided"), 
         title = replace_na(title, "None Provided"),
         area = replace_na(area, "None Provided"),
         affil_short = str_trim(str_remove_all(affiliation, "(School of.*?,)|(College of.*?,)|(Division of.*?,)|(Department of.*?,)|(, .*?Department)|(Center.*?,)")), 
         title_short = str_trim(str_remove_all(title, ",.*$")),
         area_short = str_trim(str_remove_all(area, ",.*$")))
```


## Write out the data to a CSV

```{r}
write_csv(fellows, "data/macarthur-mfd.csv")

## write_csv(data, "~/Documents/courses/visualizingsociety.com/static/data/macarthur.csv")

```

## Save the scraped webpages

There's a gotcha with objects like `bio_pages`: they cannot be straightforwardly saved to R's native data format with `save()`. The XML files are stored with external pointers to their content and cannot be "serialized" in a way that saves their content properly. If you try, when you `load()` the saved object you will get complaints about missing pointers. So instead, we'll unspool our list and  save each fellow's page individually. Then if we want to rerun this analysis without crawling everything again, we will load them in from our local saved versions using `read_html()`.

```{r}

## Get a list containing every fellow's webpage, 
## Drop the safely() error codes from the initial scrape, and 
## and also drop any NULL entries
page_list <- pluck(bio_pages, "result") %>% 
  compact()

## Make a vector of clean file names of the form "raw/macfound.org/jane_doe.html"
## One for every fellow. Same order as the page_list.
fnames <-paste0("raw/macfound.org/", 
                janitor::make_clean_names(fellows$name),
                ".html") 
names(fnames) <- fellows$name   

## Walk the elements of the page list and the file names to 
## save each HTML file under is respective clean file name
walk2(page_list, fnames, ~ write_xml(.x, file = .y))

```



## Make some plots
  
```{r}

p <- ggplot(data = fellows, 
              mapping = aes(x = year, y = age, group = year))

p + geom_jitter(alpha = 0.5, shape = 1, position = position_jitter()) +  
  geom_boxplot(outlier.size = 0, 
                   fill = my.colors("bly")[2], 
                   alpha = 0.2) + 
  theme(legend.position = "top") +
  labs(x = "Year of Award", y = "Age at Time of Award", 
       title = "Age Distribution of MacArthur Award Winners 1981-2019", 
       caption = "Kieran Healy @kjhealy socviz.co / Data: MacArthur Foundation.") + 
  scale_x_date(breaks = c(int_to_year(1981), 
                          int_to_year(1991), 
                          int_to_year(2001), 
                          int_to_year(2011), 
                          int_to_year(2019)), 
               labels = c("1981", "1991", "2001", "2011", "2019")) + 
  theme(axis.text = element_text(size = rel(1.25)),
        axis.title = element_text(size = rel(1.25)), 
        plot.title = element_text(size = rel(1.75)),
        plot.caption = element_text(size = rel(1.1)))
  

p <-   ggplot(data = subset(fellows, !is.na(sex)), 
              mapping = aes(x = year, y = age, 
                       color = sex,
                       fill = sex)) 
p + geom_jitter(alpha = 0.8, shape = 1, position = position_jitter(height = 0.2)) + 
  geom_smooth() +
  theme(legend.position = "top") +
  labs(x = "Year of Award", y = "Age at Time of Award", 
       title = "Age Distribution of MacArthur Winners over time")

fellows %>%
  group_by(sex) %>%
  tally() %>%
 mutate(pct = (n / sum(n))*100)

p <-   ggplot(data = subset(fellows, !is.na(sex)), 
              mapping = aes(x = year, y = age, color = sex, fill = sex)) 
p +   geom_boxplot(outlier.colour = NA, 
                   mapping = aes(group = interaction(year, sex), 
                                 color = sex,
                                 fill = sex), alpha = 0.3) + 
  geom_jitter(alpha = 0.4, shape = 1, 
              position = position_jitter(height = 0.2)) + 
  theme(legend.position = "top") +
  scale_fill_manual(values = my.colors("bly")) + 
  scale_color_manual(values = my.colors("bly")) +
  labs(x = "Year of Award", y = "Age at Time of Award", 
       color = "Gender", fill = "Gender",
       title = "Age Distribution of MacArthur Winners over time")


```


```{r}

fellows %>% 
  group_by(affil_short) %>%
  tally() %>%
  mutate(freq = n / sum(n),
         pct = round((freq*100), 1)) %>%
  arrange(desc(pct)) %>% 
  filter(n > 2) %>%
  select(affil_short, n, pct)


fellows %>% 
  group_by(title_short) %>%
  tally() %>%
  mutate(freq = n / sum(n),
         pct = round((freq*100), 1)) %>%
  arrange(desc(pct)) %>% 
  filter(n > 2) %>%
  select(title_short, n, pct)

fellows %>% 
  group_by(area_short) %>%
  tally() %>%
  mutate(freq = n / sum(n),
         pct = round((freq*100), 1), 
         n_lab = paste0(area_short," (", n, ")")) %>%
  arrange(desc(pct)) %>% 
  filter(pct >=1) %>%
  select(n_lab, n, pct) %>%
  ggplot(mapping = aes(x = pct, y = reorder(n_lab, pct))) + 
  geom_point(size = 3) + 
  labs(x = "Percent of All Awards", 
       y = NULL, 
       title = "MacArthur Fellowships by Area, 1981-2019", 
       subtitle = "Areas with one percent or more of all awards",
      caption = "Kieran Healy @kjhealy socviz.co / Data: MacArthur Foundation.") +
  theme(axis.text = element_text(size = rel(1.25)),
        axis.title = element_text(size = rel(1.25)), 
        plot.title = element_text(size = rel(1.75)),
        plot.caption = element_text(size = rel(1.1)))  


```



