---
title: "MacArthur Data"
author: "Kieran Healy"
date: "9/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rvest)
library(socviz)
library(showtext)
showtext_auto()
library(myriad)
import_myriad_semi()

theme_set(theme_myriad_semi())
```

# Get a List of all MacArthur Fellows 

```{r}
url <- "https://en.wikipedia.org/wiki/MacArthur_Fellows_Program"
page <- read_html(url) 
content <- html_node(page, "#mw-content-text")
years <- html_nodes(content, "h3")
winners <- html_nodes(page, ".column-width")
```

```{r}
gen_links <- function(webpage){
    # Extract the url and put the prefix back
    url_ <- webpage %>%
        rvest::html_nodes(".column-width a") %>% 
        rvest::html_attr("href") %>%
        paste0("https://en.wikipedia.org", .)
    # Extract the link text
    link_ <- webpage %>%
        rvest::html_nodes(".column-width a") %>%
        rvest::html_text() 
    tibble(name = link_, url = url_)
}

out <- gen_links(content) 

## 2018 and 2019 fn increments are missing
dummy_row <- c("[99]", "https")

## Row index of first 2019 fellow
ind <- which(out$name == "Elizabeth S. Anderson")

out <- rbind(out[1:ind-1,], dummy_row, out[ind:nrow(out),])

## Additional cites that are superfluous for year names
kill_rows <- c("[8]", "[11]", "[13]", "[14]", "[19]",
               "[47]", "[48]", "[49]", "[50]")

## Delete the superfluous rows, and then create a cumulating counter
## that tracks the award year.
out <- out %>%
  filter(name %nin% kill_rows) %>%
  mutate(year = cumsum(str_detect(name, "\\[\\d{1,2}\\]")) + 1981) %>%
  filter(!str_detect(name, "\\[\\d{1,2}\\]"))

data <- out

```

# Grab all their Wikipedia pages and put them in a list

Don't do this more than once! 

```{r, eval = FALSE}

## Grab the full Wikipedia page of every awardee. 
## Not every awardee has a page, hence safely()
bio_pages <- data$url[1:5] %>% 
  map(~ {
    message(glue::glue("* parsing: {.x}"))
    Sys.sleep(5) # try to be polite
    safely(read_html)(.x)
  })

## Name the list elements
bio_pages <- set_names(bio_pages, data$name)
```

There's a gotcha with objects like `bio_pages`: they are _serialized_, which means more or less that they contain pointers to data stored live in memory (in this case the HTML files for each awardee). The catch is that serialized objects cannot be saved to disk in a way that preserves their contents. Normally we can `save()` data objects and `load()` them later, but XML/rvest objects can't be written straight to disk. So instead, we'll grab each page and save it individually. Then if we want to rerun this analysis without crawling all the pages again, we will load them in from our local saves.

```{r}
## Get clean file names
fnames <-paste0("raw/", janitor::make_clean_names(data$name),".html") 
names(fnames) <- data$name   

## Get every webpage, dropping the safely() error codes
page_list <- pluck(bio_pages, "result") 

## Get a vector of null names -- i.e. awardees who have no wikipedia page
null_names <- keep(page_list, is.null)

## Drop the empty entries from the list of pages to save
page_list <- compact(page_list)

## Remove null names from our list of filenames to save
ind <- names(fnames) %nin% names(null_names)
fnames <- fnames[ind]

## Walk the page_list, saving each HTML file with its respective clean file name
walk2(page_list, fnames, ~ write_xml(.x, file = .y))
```

# Parse the pages, searching for birth years

Two main options: look for dates in the main text (a pain; and error prone); extract `bday` element from the `vcard` (accurate, but mostly missing). Here are functions to do each of these.

```{r}

## Born 1950 / born 1950
wiki_born <- "[Bb]orn\\s+\\d{4}"

## January 10, 1950
us_date <- "(Jan(?:uary)?|Feb(?:ruary)?|Mar(?:ch)?|Apr(?:il)?|May|Jun(?:e)?|Jul(?:y)?|Aug(?:ust)?|Sep(?:tember)?|Oct(?:ober)?|Nov(?:ember)?|Dec(?:ember)?)\\s+(\\d{1,2}),\\s+(\\d{4})"

## 10 January 1950
uk_date <- "(\\d{1,2})\\s+(Jan(?:uary)?|Feb(?:ruary)?|Mar(?:ch)?|Apr(?:il)?|May|Jun(?:e)?|Jul(?:y)?|Aug(?:ust)?|Sep(?:tember)?|Oct(?:ober)?|Nov(?:ember)?|Dec(?:ember)?)\\s+(\\d{4})"

## Each of the above
full_reg <- paste0("(", wiki_born, ")|(", us_date, ")|(", uk_date, ")")

body_age_string <- function(x, reg = full_reg) {
    ifelse(is.null(x), return(NA), FALSE)
    out <- rvest::html_node(x, "#mw-content-text") %>%
            rvest::html_text() %>%
            str_remove("in \\d{4}") %>%
            str_extract(reg) %>%
            str_extract("\\d{4}") %>%
            as.integer()
    
    ifelse(identical(out, character(0)), NA, out)
}

bday_age_string <- function(x) {
    ifelse(is.null(x), return(NA), FALSE)
      
    out <- rvest:::html_nodes(x, ".bday") %>%
      rvest::html_text() %>%
       str_extract("\\d{4}") %>%
       as.integer()
    ifelse(identical(out, character(0)), NA, out)
}


```

```{r}
  
body_out <- bio_pages %>% 
  map("result") %>%
  map_int(body_age_string)

body_out

bday_out <- bio_pages %>% 
  map("result") %>%
  map_int(bday_age_string)

bday_out

```

# Bind the ages in to the data.

```{r}

data <- out

data <- data %>% add_column(body_out)
data <- data %>% add_column(bday_out)

data$born <- data$bday_out
ind <- is.na(data$born)

data$born[ind] <- data$body_out[ind]

```

# Clean Errors and Problem Cases

E.g., 

```{r}
range(data$born, na.rm = TRUE)

data %>% filter(born == 1675)

data %>% filter(name == "Mark S. Wrighton")

```

Or,

```{r}
bad_cases <- data %>% filter(age < 20) 

bad_cases
```

```{r}

## Ad Hoc birth year fixes go here.
data <- data %>%
  mutate(born_rc = case_when(
    name == "Mark S. Wrighton" ~ as.integer(1949), 
                             name == "Wampanoag" ~ NA_integer_,
                             TRUE ~ born))

data %>% filter(year == 2018, is.na(age)) %>% select(name)

data <- data %>% 
  mutate(age = year - born_rc, 
         age = replace(age, age < 20, NA)) 

## General ad hoc fixes
data <- data %>%
  filter(name %nin% c("elder rights", "Alutiiq", "Benetech"))
## Problem parsing at Marie-Therese Connolly (Elder Rights is her job but it ended up as a name)

## Ad Hoc age fixes go here.
## Do it by award year for clarity, I guess
## These age-at-award values were taken directly from macfound.org

## 2019 class
## data %>% filter(year == 2019, is.na(age)) %>% select(name)
data <- data %>%
  mutate(age = case_when(
    name == "Sujatha Baliga" ~ as.integer(48), 
    name == "Danielle Citron" ~ as.integer(50), 
    name == "Lisa Daugaard" ~ as.integer(53),    
    name == "Lisa Daugaard" ~ as.integer(53),    
    name == "Annie Dorsen" ~ as.integer(45),
    name == "Andrea Dutton" ~ as.integer(46),
    name == "Mary Halvorson" ~ as.integer(38),
    name == "Sadiya Hartman" ~ as.integer(58),
    name == "Stacy Jupiter" ~ as.integer(43),
    name == "Zachary Lippman" ~ as.integer(41),    
    name == "Kelly Lytle Hernandez" ~ as.integer(45),
    name == "Jeffrey Alan Miller" ~ as.integer(35),
    name == "Jerry X. Mitrovica" ~ as.integer(58),
    name == "Emmanual Pratt" ~ as.integer(42),
    name == "Cameron Rowland" ~ as.integer(39),
    name == "Vanessa Ruta" ~ as.integer(45),
    name == "Joshua Tenenbaum" ~ as.integer(47),    
    name == "Jenny Tung" ~ as.integer(37),
    TRUE ~ as.integer(age)))

## 2018 class 
## data %>% filter(year == 2018, is.na(age)) %>% select(name)
data <- data %>%
  mutate(age = case_when(
  name == "Clifford Brangwynne" ~ as.integer(40),
  name == "Natalie Diaz" ~ as.integer(40),
  name == "Livia S. Eberlin" ~ as.integer(32),
  name == "Gregg Gonsalves" ~ as.integer(54),
  name == "Becca Heller" ~ as.integer(36),
  name == "Raj Jayadev" ~ as.integer(43),
  name == "Titus Kaphar" ~ as.integer(42),
  name == "Kristina Olson" ~ as.integer(37),
  name == "Lisa Parks" ~ as.integer(51),
  name == "Rebecca Sandefur" ~ as.integer(47),
  name == "Allan Sly" ~ as.integer(36),
  name == "Sarah T. Stewart-Mukhopadhyay" ~ as.integer(45),
  name == "Doris Tsao" ~ as.integer(42),
  name == "Ken Ward Jr" ~ as.integer(50),  
  TRUE ~ as.integer(age)))


## 2017 class 
## data %>% filter(year == 2017, is.na(age)) %>% select(name)
data <- data %>%
  mutate(age = case_when(
  name == "Sunil Amrith" ~ as.integer(38),
  name == "Greg Asbed" ~ as.integer(54),
  name == "Jason De León" ~ as.integer(40),
  name == "Rami Nashashibi" ~ as.integer(45),
  name == "Kate Orff" ~ as.integer(45),
  name == "Betsy Levy Paluck" ~ as.integer(39),
  name == "Yuval Sharon" ~ as.integer(37),
  name == "Gabriel Victora  " ~ as.integer(40),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2016, is.na(age)) %>% select(name)
data <- data %>%
  mutate(age = case_when(
    name == "Ahilan Arulanantham" ~ as.integer(43),
    name == "Daryl Baldwin" ~ as.integer(53),
    name == "Anne Basting" ~ as.integer(51),
    name == "Josh Kun" ~ as.integer(45),
    name == "Dianne Newman" ~ as.integer(44),
    name == "Victoria Orphan" ~ as.integer(44),
    name == "Manu Prakash" ~ as.integer(36),
    name == "José A. Quiñonez" ~ as.integer(45),
    name == "Lauren Redniss" ~ as.integer(42),
    name == "Sarah Stillman" ~ as.integer(32),
    name == "Bill Thies" ~ as.integer(38),
    name == "Jin-Quan Yu" ~ as.integer(50),
    TRUE ~ as.integer(age)))
    
## data %>% filter(year == 2015, is.na(age)) %>% select(name)    
## Weird: who and where is Lee Curle?

data <- data %>%
  mutate(age = case_when(
  name == "Kartik Chandran" ~ as.integer(41),
  name == "Gary Cohen" ~ as.integer(59),
  name == "Matthew Desmond" ~ as.integer(35),
  name == "John Novembre" ~ as.integer(37),
  name == "Christopher Ré" ~ as.integer(36),
  name == "Marina Rustow" ~ as.integer(46),
  name == "Juan Salgado" ~ as.integer(46),
  name == "Alex Truesdell" ~ as.integer(59),
  name == "Basil Twist" ~ as.integer(46),
  TRUE ~ as.integer(age)))

## data %>% filter(year == 2014, is.na(age)) %>% select(name)    

data <- data %>%
  mutate(age = case_when(
  name == "Danielle Bassett" ~ as.integer(32),
  name == "Craig Gentry" ~ as.integer(41),
  name == "John Henneberger" ~ as.integer(59),
  name == "Jonathan Rapping" ~ as.integer(48),
  TRUE ~ as.integer(age)))

## data %>% filter(year == 2013, is.na(age)) %>% select(name)    
## No data for Nirenberg

data <- data %>%
  mutate(age = case_when(
  name == "C. Kevin Boyce" ~ as.integer(39),
  name == "Jeffrey Brenner" ~ as.integer(44),
  name == "Craig Fennie" ~ as.integer(40),
  name == "Robin Fleming" ~ as.integer(57),
  name == "Carl Haber" ~ as.integer(54),
  name == "David Lobell" ~ as.integer(34),
  name == "Ana Maria Rey" ~ as.integer(36),
  TRUE ~ as.integer(age)))

## data %>% filter(year == 2012, is.na(age)) %>% select(name)    

data <- data %>%
  mutate(age = case_when(
  name == "Natalia Almada" ~ as.integer(37),
  name == "Elissa Hallem" ~ as.integer(34),
  name == "Maurice Lim Miller" ~ as.integer(66),
  name == "Dylan C. Penningroth" ~ as.integer(41),
  name == "Nancy Rabalais" ~ as.integer(62),
  name == "Melody Swartz" ~ as.integer(43),
  name == "Benjamin Warf" ~ as.integer(54),  
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2011, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Marie-Therese Connolly" ~ as.integer(54),
  name == "Tiya Miles" ~ as.integer(41),
  name == "Matthew Nock" ~ as.integer(38),
  TRUE ~ as.integer(age)))

## data %>% filter(year == 2010, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Kelly Benoit-Bird" ~ as.integer(34),
  name == "Carlos D. Bustamante" ~ as.integer(35),
  name == "Shannon Lee Dawdy" ~ as.integer(43),
  name == "Nergis Mavalvala" ~ as.integer(42),
  name == "Sebastian Ruth" ~ as.integer(35),
  name == "Dawn Song" ~ as.integer(35),
  name == "Elizabeth Turk" ~ as.integer(48),
  TRUE ~ as.integer(age)))

## data %>% filter(year == 2009, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Timothy Barrett" ~ as.integer(59),
  name == "Lin He" ~ as.integer(35),
  name == "James Longley" ~ as.integer(37),
  name == "L. Mahadevan" ~ as.integer(44),
  name == "Elyn Saks" ~ as.integer(53),
  name == "Jill Seaman" ~ as.integer(57),
  name == "Daniel Sigman" ~ as.integer(40),
  name == "Mary Tinetti" ~ as.integer(58),
  name == "Theodore Zoli" ~ as.integer(43),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2008, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Kirsten Bomblies" ~ as.integer(34),
  name == "Susan Mango" ~ as.integer(47),
  name == "David R. Montgomery" ~ as.integer(46),
  name == "Sally Temple" ~ as.integer(49),
  name == "Rachel Wilson" ~ as.integer(34),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2007, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Peter Cole" ~ as.integer(50),
  name == "Marc Edwards" ~ as.integer(43),
  name == "Michael Elowitz" ~ as.integer(37),
  name == "Cheryl Hayashi" ~ as.integer(40),
  name == "Yoky Matsuoka" ~ as.integer(36),
  name == "Paul Rothemund" ~ as.integer(35),
  name == "Shen Wei" ~ as.integer(39),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2006, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Lisa Curran" ~ as.integer(45),
  name == "Victoria Hale" ~ as.integer(45),
  name == "Adrian Nicole LeBlanc" ~ as.integer(42),
  name == "D. Holmes Morton" ~ as.integer(55),
  name == "John A. Rich" ~ as.integer(47),
  name == "Matias Zaldarriaga" ~ as.integer(35),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2005, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Edet Belzberg" ~ as.integer(35),
  name == "Lu Chen" ~ as.integer(33),
  name == "Michael Cohen" ~ as.integer(61),
  name == "Joseph Curtin" ~ as.integer(52),
  name == "Claire Gmachl" ~ as.integer(38),
  name == "Michael Manga" ~ as.integer(37),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2004, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Angela Belcher" ~ as.integer(37),
  name == "Gretchen Berland" ~ as.integer(40),
  name == "Joseph DeRisi" ~ as.integer(35),
  name == "Katherine Gottlieb" ~ as.integer(52),
  name == "John Kamm" ~ as.integer(53),
  name == "Naomi Leonard" ~ as.integer(40),
  name == "Vamsi Mootha" ~ as.integer(33),
  name == "Cheryl Rogowski" ~ as.integer(43),
  TRUE ~ as.integer(age)))

## data %>% filter(year == 2003, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Sarah H. Kagan" ~ as.integer(41),
  name == "Ned Kahn" ~ as.integer(43),
  name == "Nawal M. Nour" ~ as.integer(37),
  name == "Amy Rosenzweig" ~ as.integer(36),
  name == "Eve Troutt Powell" ~ as.integer(42),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2002, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
    name == "David B. Goldstein" ~ as.integer(51),
    name == "Janine Jagger" ~ as.integer(52),
    name == "Daniel Jurafsky" ~ as.integer(39),
    name == "Erik Mueggler" ~ as.integer(40),
    name == "Sendhil Mullainathan" ~ as.integer(29),
    name == "Lee Ann Newsom" ~ as.integer(45),
    name == "Daniela L. Rus" ~ as.integer(39),
    name == "Brian Tucker" ~ as.integer(56),
    name == "Paul Wennberg" ~ as.integer(40),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2001, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Christopher Chyba" ~ as.integer(41),
  name == "Brooks Pate" ~ as.integer(36),
  name == "Julie Su" ~ as.integer(32),
  name == "David Wilson" ~ as.integer(55),
  TRUE ~ as.integer(age)))


## data %>% filter(year == 2000, is.na(age)) %>% select(name)    
data <- data %>%
  mutate(age = case_when(
  name == "Susan E. Alcock" ~ as.integer(39),
  name == "K. Christopher Beard" ~ as.integer(38),
  name == "Lucy Blake" ~ as.integer(42),
  name == "Laura Otis" ~ as.integer(39),
  name == "Susan E. Sygall" ~ as.integer(47),
  name == "Gina G. Turrigiano" ~ as.integer(37),
  TRUE ~ as.integer(age)))


```

## Write out the data to a CSV

```{r}
write_csv(data, "data/macarthur.csv")

## write_csv(data, "~/Documents/courses/visualizingsociety.com/static/data/macarthur.csv")

```



## Make some plots
  
```{r}

data %>%
  ggplot(mapping = aes(x = socviz::int_to_year(year), 
                       y = age)) + 
  geom_jitter(alpha = 0.8, shape = 1, position = position_jitter(height = 0.2)) + 
  geom_smooth() +
  labs(x = "Year of Award", y = "Age at Time of Award", 
       title = "Age Distribution of MacArthur Winners", 
       subtitle = "Incomplete data, please don't share")

```

